Bayesian Exploration in Deep Reinforcement Learning

### 2022-01-27 papers/main.tex
TEMPORARY DEPLOY COMMIT

\title{Parameter-Free Exploration in Deep Reinforcement Learning}

@@

The difficult part in \eqref{eq:fkl} is to estimate \(\nabla_\f \log q(\f^\X)\)
and \(\nabla_\f \log p(\f^\X)\). \(\nabla_\f \log q(\f^\X)\) is likely 
intractable, considering \(q_\phi\) is a neural network with stochastic weights.
Depending on how we define the prior, however, \(\nabla_\f \log p(\f^\X)\)
can be easy to compute analytically. To reduce variance in the gradients we
define tractable priors in this paper. To estimate the log-density gradient
\(\nabla_\f \log q(\f^\X)\), they use a spectral Stein gradient estimator
\citep{shi_spectral_2018}.

@@

We can use the functional variational bayesian neural
network (FVBNN)~\citep{sun_functional_2019} framework discussed earlier.
NoisyNet uses one sample from Q for each optimization step. We will instead use
\(k\) samples from Q. This lets us add the KL-term from FVBNN to the loss function that
encourages stochasticity.
The maximization target becomes:
\begin{equation}
    \log p(\D_\omega^N \vert f) - \KL \left[ q(\f^{\D_\omega^N}, \f^\mathbf{M}) \Vert p(\f^{\D_\omega^N}, \f^\mathbf{M}) \right],
\end{equation}
where \(\f^{\D_\omega^N}\) is \(f\) applied to the dataset \(\D_\omega^N\), and \(\f^\mathbf{M}\) is \(f\) applied to
a set of i.i.d. random points \(\mathbf{M} = \{m_i \sim c \mid i=1,\dots,k\} \subseteq \state \times \action\)
with full support. We will let the measurement \(\mathbf{M}\) be the the set of all state-action
pairs for states we have already explored. If the MDP is ergodic, then \(\textbf{M}\) will eventually
have full support in \(\state \times \action\).

@@

@@

\section{Experiments}
Our method most closely resembles that of NoisyNet~\cite{fortunato_noisy_2019}, so
all experiments will mainly compare the performance of these two methods.
\subsection{Uninformative Prior}
First, it is interesting to see how our method will compare to NoisyNet
in environments without the advantage of prior knowledge.
We hypothesize that the added KL-term will push the algorithm to keep exploring unexplored
state-action pairs more than NoisyNet, thus

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{charts/Cartpole_score}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{charts/MountainCar_score}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{charts/Cartpole_score}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{charts/MountainCar_score}
    \end{subfigure}
    \caption{Example figure 1}
    \label{fig:fig_1}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{example-image}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{example-image}
    \end{subfigure}
    \includegraphics[width=\columnwidth]{example-image}
    \caption{Performance on the CartPole-v1 environment.}
\end{figure}

@@

\section{Conclusion}

>>>>

\title{Bayesian Exploration in Deep Reinforcement Learning}

@@

The first term is just the expected value of the score function, which is zero.
The difficult part in \eqref{eq:fkl} is to estimate \(\nabla_\f \log q(\f^\X)\)
and \(\nabla_\f \log p(\f^\X)\). \(\nabla_\f \log q(\f^\X)\) is likely 
intractable, considering \(q_\phi\) is a neural network with stochastic weights.
Depending on how we define the prior, however, \(\nabla_\f \log p(\f^\X)\)
can be easy to compute analytically. To reduce variance in the gradients we
define tractable priors in this paper. To estimate the log-density gradient
\(\nabla_\f \log q(\f^\X)\), they use the Spectral Stein Gradient Estimator (SSGE)
\citep{shi_spectral_2018}. In short, this is a method for estimating the gradient
function of implicit distributions using approximations to eigenfunctions of a 
kernel based operator. We bring three hyperparameters into the algorithm from this
method, the number of samples \(k\) from the implicit distribution used to approximate
the gradient, the number of eigenvectors \(J\) used to approximate the gradient, and
\(\eta\), a regularization parameter that smooths the gradient function.

The full minimization target for the functional Bayesian neural network becomes
\begin{multline}
    {\frac{1}{\lvert \D_s \rvert}} \sum_{()\bm{x}, y) \in D_s} \E_{q_\phi}
    \left[ p(y \vert f(\bm{x})) \right]\\ - \lambda \KL
    \left[ q(\f^{\D_s}, \f^\mathbf{M}) \Vert p(\f^{\D_s}, \f^\mathbf{M}) \right].
\end{multline}

In order for this to match the functional ELBO and be a proper lower bound
for \(\log p(\D)\), \(\lambda\) should be set as \(\frac{1}{\lvert \D \rvert}\).
\cite{sun_functional_2019} note, however, that the minimization target is only a lower
bound for \(\KL \left[ \lvert q(f) \Vert p(f) \right]\), so a larger
value for \(\lambda\) is likely necessary to prevent overfitting. They use,
\(\lambda = \frac{1}{\lvert D_s \rvert}\), a value of one over the batch size instead.
This means that the minimization target becomes a proper lower bound of \(\log p(\D_s)\).

@@

We can use the functional variational bayesian neural
network (FVBNN)~\citep{sun_functional_2019} framework discussed earlier.
NoisyNet uses one sample from Q for each optimization step. We will instead use
\(k\) samples from Q. This lets us add the KL-term from FVBNN to the loss function that
encourages stochasticity.
The maximization target becomes:
\begin{equation}
    \log p(\D_\omega^N \vert f) - \lambda \KL \left[ q(\f^{\D_\omega^N}, \f^\mathbf{M}) \Vert p(\f^{\D_\omega^N}, \f^\mathbf{M}) \right],
\end{equation}
where \(\f^{\D_\omega^N}\) is \(f\) applied to the dataset \(\D_\omega^N\), and \(\f^\mathbf{M}\) is \(f\) applied to
a set of i.i.d. random points \(\mathbf{M} = \{m_i \sim c \mid i=1,\dots,k\} \subseteq \state \times \action\)
with full support. We will let the measurement \(\mathbf{M}\) be the the set of all state-action
pairs for states we have already explored. If the MDP is ergodic, then \(\textbf{M}\) will eventually
have full support in \(\state \times \action\).

@@

To calculate \(\log p(\D_\omega^N \vert f)\) we assume that the observation noise is sampled
from a Gaussian distribution. We then model the network to produce two output vectors,
\(\mu\) and \(\rho = \log \sigma\). This gives the following loss function:

\begin{equation}
    \mathcal{L} = \frac{1}{\lvert \D_s \rvert}\sum_{i=1}^{\lvert \D_s \rvert} \rho_i + \left( \mu_i - q_i \right)^2
    \exp (-\rho_i) + \lambda \KL[q \Vert p].
\end{equation}

We take inspiration from \cite{sun_functional_2019} and let \(\lambda\) be one over
the batch size, \(\frac{1}{\lvert \D_s \rvert}\).

@@

\section{Experiments}
Our method most closely resembles that of NoisyNet~\cite{fortunato_noisy_2019}, so
all experiments will mainly compare the performance of these two methods.

\subsection{Details and Hyperparameters}
We compare our method with NoisyNet~\citep{fortunato_noisy_2019} on four different
environments in the OpenAI Gym~\citep{brockman_openai_2016-2}. The hyperparameters
are the same as those described in \cite{han_nrowan-dqn_2020}. Our method additionally
has hyperparameters related to the functional KL-divergence. These are shown in
Table~\ref{tab:hyperparams}.

\begin{table}
    \centering
    \caption{Hyperparameters for LFM.}\label{tab:hyperparams}
    \begin{tabular}{rl}
      \toprule % from booktabs package
      Hyperparameter & Value \\
      \midrule % from booktabs package
      \(k\) (number of samples) & 100\\
      \(J\) (Number of eigenvalues) & 6\\
      \(\lambda\) & \(\frac{1}{\lvert \mathcal{D}_s \rvert}\)\\
      \(\eta\) (Regularization coefficient for SSGE) & 0.95\\
      \bottomrule % from booktabs package
    \end{tabular}
\end{table}

\subsection{Uninformative Prior}
First, it is interesting to see how our method will compare to NoisyNet
in environments without the advantage of prior knowledge. The training curves
of our model compared to that of standard NoisyNet-DQN on the four selected OpenAI
Gym~\citep{brockman_openai_2016} enivornments is shown in Figure~\ref{fig:training_curves}.
We hypothesize that the added KL-term will push the algorithm keep weight uncertinaty high
in order to keep a higher uncertainty in the output. While both methods are able to
find an optimal policy in Cartpole, our method method uses considerably fewer frames
to do so.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{charts/Cartpole_score}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{charts/MountainCar_score}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{example-image}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{example-image}
    \end{subfigure}
    \caption{Comparison of learning curves of LFM and NoisyNet-DQN on
    Pong, Cartpole, MountainCar, and Acrobot averaged over 100 episodes.}
    \label{fig:training_curves}
\end{figure}

\cite{fortunato_noisy_2019} noted that the learned variance in their weight increased
in some games despite there existing an optimal deterministic solution, and the loss
giving no incentive to maintain any uncertainty. Figure~\ref{fig:mean_std} shows
the mean-absolute standard deviation for the penultimate and final layer. It is 
interesting to see that the standard deviation for NoisyNet continous to decrease
throughout the training, while LFM decreases faster, but then flattens out earlier.
This seems to indicate that it has found a policy that

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{charts/penult}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{charts/ult}
    \end{subfigure}
    \caption{Comparison of mean-absolution standard deviation for NoisyNet and LFM
    on Cartpole.}
    \label{fig:mean_std}
\end{figure}

@@

% \begin{figure}
%     \centering
%     \begin{subfigure}[b]{0.45\columnwidth}
%         \centering
%         \includegraphics[width=\columnwidth]{example-image}
%     \end{subfigure}
%     \begin{subfigure}[b]{0.45\columnwidth}
%         \centering
%         \includegraphics[width=\columnwidth]{example-image}
%     \end{subfigure}
%     \includegraphics[width=\columnwidth]{example-image}
%     \caption{Performance on the CartPole-v1 environment.}
% \end{figure}

@@

\section{Conclusion}
We have developed a method for exploration with a DQN-style algorithm that can
effectively utilize domain knowledge for faster learning.
